Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
For real now
Major haircut for the intro -- feels much better now!

For the confidentiality and integrity properties, we consider both {\em
  stepwise} and {\em extensional} variants.  The simpler {stepwise} variants
formalize the intuition that a caller's local data is {\em never} read or
modified during a call. The extensional variants allow a callee to read from
from and write to their caller's stack frame, as long as nothing that they
read affects their observable behavior and nothing they write affects the
observable behavior of the caller after they return.  \bcp{Please check: is
  that accurate?}

@@

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)

We test against their {\em Depth Isolation} micro-policy, in which memory cells within each stack
frame are tagged with the identity of the function activation
that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. This is simple and corresponds
to intuitive notions of stack safety, and we argue that by describing its behavior
we approach a reasonable stack-safety property that other micro-policies can be
validated against.

% Randomized testing is a relatively low-cost way to validate a potential enforcement
% mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
% of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
% covers potentially erroneous programs, we test intentionally buggy policies alongside
% the real one.

A second challenge is that enforcement mechanisms may weaken their
protections\bcp{``weaken their protections'' doesn't convey what we mean
  very well}
in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
\citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
ways of identifying functions and deciding exactly which accesses are allowed.
These vary in their efficiency and in the spatial and temporal precision of their
security guarantees.
The Depth Isolation policy prevents a callee from writing outside its frame
at all, which is simple and intuitively correct, but it must manually clear the tags
on the frame on each return, which is very expensive. Conversely, their
\emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. This violates our
straightforward properties, but ntuitively,
this form of laziness is justified if any dangerous writes will
ultimately be detected before they can affect program behavior.

\bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
What does ``affecting program behavior'' really mean? This micro-policy
calls for a second characterization of stack safety in terms of the externally
observable behavior of a stack-safe system. Such a characterization will have the
additional benefit of being more {\em extensional}, concerned primarily with the
external behavior of the system. In contract, our stepwise properties are {\em intensional},
focusing on internal details that may not actually matter to overall security.

\bcp{Ditto, even moreso.}
To formalize this intuition, we develop an \emph{observational integrity} property which
characterizes enforcement mechanisms that detect violations between the point when
they occur and the point when they make an observable impact. Intuitively this property
can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
integrity implies observational integrity. \sna{TODO: check this.}

We additionally sketch an associated {\em observational confidentiality} property
makes confidentiality extensional in the same way, in order to form an
{\em observational stack safety} that is highly extensional.
We argue that, despite its somewhat greater complexity,
it may be a better specification for
stack safety enforcement mechanisms in general.

%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
As it turns out, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy enforces observational integrity, but not confidentiality,
whether stepwise or observational.
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality occurs in corner cases where callees
can leak data between one another.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
confidentiality, albeit at some performance cost.  \bcp{Did Nick and Andre
  know about this deficiency?  Did they mention it in their paper?  (If
  neither, we should make a bigger deal out of the fact that we discovered it.)}

Our initial description of these properties assumes a single simple stack with no sharing
between callers and callees: all parameters and return values are passed in registers.  Later,
we show how to refine it to allow passing of scalar stack data, granting a
callee access to the caller's data. We do not model more sophisticated sharing,
such as taking the address of a variable and treating it like a pointer.

Finally, we define integrity and confidentiality properties for a coroutine system featuring
a static layout, in which each coroutine's stack is constrained to a fixed region of memory.
Though simplistic, this extension is a step toward concurrency and demonstrates that
our approach is flexible enough to handle more sophisticated styles of
control flow.

>>>>

For the confidentiality and integrity properties, we consider both {\em
  stepwise} and {\em observational} variants.  The simpler {stepwise}
variants formalize the intuition that a caller's local data is {\em never}
read or modified during a call. The more extensional observational variants
allow a callee to read from from and write to their caller's stack frame, as
long as nothing that they read affects their observable behavior and nothing
they write affects the observable behavior of the caller after they return.
\bcp{Please check: is that accurate?}

@@

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)
%
We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then controlled by the activation that is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but testing
reveals that it does not enforce confidentiality (either stepwise or
observational): it violates of the temporal aspect of confidentiality in
corner cases where data can leak across calls to the same callee.  We
propose a variant of Lazy Tagging and Clearing that \emph{does} enforce
confidentiality (as far as our testing can tell), albeit at some performance
cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)}

% A second challenge is that enforcement mechanisms may weaken their
% protections\bcp{``weaken their protections'' doesn't convey what we mean
%   very well}
% in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
% \citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
% ways of identifying functions and deciding exactly which accesses are allowed.
% These vary in their efficiency and in the spatial and temporal precision of their
% security guarantees.
% The Depth Isolation policy prevents a callee from writing outside its frame
% at all, which is simple and intuitively correct, but it must manually clear the tags
% on the frame on each return, which is very expensive. Conversely, their
% \emph{Lazy Tagging and Clearing} policy improves
% performance by assigning locations to the active function when they are written, which
% permits  ``illicit'' writes to inaccessible stack frames so long as the
% written locations are not subsequently read by their ``real'' owners. This violates our
% straightforward properties, but ntuitively,
% this form of laziness is justified if any dangerous writes will
% ultimately be detected before they can affect program behavior.

% \bcp{This discussion is getting too long for the intro.  Could it be trimmed.}
% What does ``affecting program behavior'' really mean? This micro-policy
% calls for a second characterization of stack safety in terms of the externally
% observable behavior of a stack-safe system. Such a characterization will have the
% additional benefit of being more {\em extensional}, concerned primarily with the
% external behavior of the system. In contract, our stepwise properties are {\em intensional},
% focusing on internal details that may not actually matter to overall security.

% \bcp{Ditto, even moreso.}
% To formalize this intuition, we develop an \emph{observational integrity} property which
% characterizes enforcement mechanisms that detect violations between the point when
% they occur and the point when they make an observable impact. Intuitively this property
% can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
% integrity implies observational integrity. \sna{TODO: check this.}

% We additionally sketch an associated {\em observational confidentiality} property
% makes confidentiality extensional in the same way, in order to form an
% {\em observational stack safety} that is highly extensional.
% We argue that, despite its somewhat greater complexity,
% it may be a better specification for
% stack safety enforcement mechanisms in general.

%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}

Our first presentation of these properties assumes a single simple stack
with no sharing between callers and callees: all parameters and return
values are passed in registers.  Later, we show how to refine it to allow
passing of scalar stack data, granting a callee access to the caller's
data. We do not model more sophisticated sharing, such as taking the address
of a variable and treating it like a pointer.
%
Finally, we define integrity and confidentiality properties for a coroutine
system featuring a static layout, in which each coroutine's stack is
constrained to a fixed region of memory.  Though simplistic, this extension
is a step toward concurrency and demonstrates that our approach is flexible
enough to handle more sophisticated styles of control flow.
